The Power of Two Random Choices\\ \large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Tr\"aff and Francesco Versaci, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}

### 2014-03-11 article/article.tex
add comma after w.h.p. if sentence continues

\subsection{Single-choice BiB}
\label{sec:single-choice}
The \emph{Single-choice} strategy is a simple allocation scheme for the BiB problem and chooses one bin uniformly and independently at random for each ball. Using this strategy, the maximum load of the fullest bin will be
\begin{comment}
\footnote{A more accurate result exists: The maximum load is $\load = \Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$~\cite{G91}.}
\end{comment}
\begin{align*}
\load = \frac{\ln n}{\ln \ln n}  \left(1 + o\left(1\right)\right)
\end{align*}
for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E_n$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E_n \right] \geq 1 - n^{-\alpha}$ for some arbitrary constant $\alpha > 0$.}~\cite{RS98}. This implies (see Theorem \ref{theorem:asymptotic} in Appendix \ref{sec:asymptotic})
\begin{align*}
\load\in \Theta\left(\frac{\ln n}{\ln \ln n}\right).
\end{align*}
Note that, since the average load is 1, $\delta \in \Theta\left( L_{max} \right)$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is ~\cite{RS98}
\begin{align*}
\delta = \Theta\left(\sqrt{\frac{m  \ln n}{n}}\right).
\end{align*}

@@

To maximize the throughput of the servers, a uniform distribution of tasks to servers is desirable. While a central dispatcher can easily achieve this, it also poses a bottleneck within a distributed system, since querying $m$ servers for their current load involves significant overhead (e.g., sending a message to each server and waiting for the reply). As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins). Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most $O\left(\ln \ln n\right)$, w.h.p.,~\cite{KLM92,RS98, MRS01}.

@@

\begin{theorem}
\label{theorem:greedy}
Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is at most
\begin{align*}
\load = \frac{\ln \ln n}{\ln d}+ O(1)
\end{align*}
w.h.p.~for all $t$~\cite{ABKU99, BCSV06}.
\end{theorem}

@@

\begin{theorem}
\label{theorem:aglm}
Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Always-go-left algorithm, the maximum load is at most 
\begin{align*}
\load = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
\end{align*}
w.h.p.~for all $t$~\cite{VOC03}.
\end{theorem}

@@

\section{Conclusion}
\label{sec:conclusion}
BiB games have many applications in Computer Science and have been studied extensively. The single-choice BiB strategy chooses a bin uniformly and independently at random for each ball. The classical result for this scheme upper-bounds the gap with~\cite{RS98}
\begin{align*}
\delta =  \Theta\left(\sqrt{\frac{m  \log n}{n}}\right).
\end{align*}
w.h.p.
In this paper we have presented two different multiple-choice BiB placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a gap of 
\begin{align*}
\delta = \frac{\ln\ln n}{\ln d} + O(1)
\end{align*}
w.h.p.~and the Always-go-left scheme due to V\"ocking~\cite{VOC03}, which achieves a slightly better gap of
\begin{align*}
\delta = \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
\end{align*}
w.h.p.~for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds for the case $m=n$ in detail.

>>>>

\subsection{Single-choice BiB}
\label{sec:single-choice}
The \emph{Single-choice} strategy is a simple allocation scheme for the BiB problem and chooses one bin uniformly and independently at random for each ball. Using this strategy, the maximum load of the fullest bin will be
\begin{comment}
\footnote{A more accurate result exists: The maximum load is $\load = \Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$~\cite{G91}.}
\end{comment}
\begin{align*}
\load = \frac{\ln n}{\ln \ln n}  \left(1 + o\left(1\right)\right)
\end{align*}
for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E_n$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E_n \right] \geq 1 - n^{-\alpha}$ for some arbitrary constant $\alpha > 0$.}~\cite{RS98}. This implies (see Theorem \ref{theorem:asymptotic} in Appendix \ref{sec:asymptotic})
\begin{align*}
\load\in \Theta\left(\frac{\ln n}{\ln \ln n}\right).
\end{align*}
Note that, since the average load is 1, $\delta \in \Theta\left( L_{max} \right)$ in this case. For the heavily loaded case, w.h.p., the maximum gap is ~\cite{RS98}
\begin{align*}
\delta = \Theta\left(\sqrt{\frac{m  \ln n}{n}}\right).
\end{align*}

@@

To maximize the throughput of the servers, a uniform distribution of tasks to servers is desirable. While a central dispatcher can easily achieve this, it also poses a bottleneck within a distributed system, since querying $m$ servers for their current load involves significant overhead (e.g., sending a message to each server and waiting for the reply). As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins). Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most $O\left(\ln \ln n\right)$, w.h.p., \cite{KLM92,RS98, MRS01}.

@@

\begin{theorem}
\label{theorem:greedy}
Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is at most
\begin{align*}
\load = \frac{\ln \ln n}{\ln d}+ O(1)
\end{align*}
w.h.p., for all $t$~\cite{ABKU99, BCSV06}.
\end{theorem}

@@

\begin{theorem}
\label{theorem:aglm}
Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Always-go-left algorithm, the maximum load is at most 
\begin{align*}
\load = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
\end{align*}
w.h.p.,~for all $t$~\cite{VOC03}.
\end{theorem}

@@

\section{Conclusion}
\label{sec:conclusion}
BiB games have many applications in Computer Science and have been studied extensively. The single-choice BiB strategy chooses a bin uniformly and independently at random for each ball. The classical result for this scheme upper-bounds the gap with~\cite{RS98}
\begin{align*}
\delta =  \Theta\left(\sqrt{\frac{m  \log n}{n}}\right).
\end{align*}
w.h.p.
In this paper we have presented two different multiple-choice BiB placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a gap of 
\begin{align*}
\delta = \frac{\ln\ln n}{\ln d} + O(1)
\end{align*}
w.h.p., and the Always-go-left scheme due to V\"ocking~\cite{VOC03}, which achieves a slightly better gap of
\begin{align*}
\delta = \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
\end{align*}
w.h.p., for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds for the case $m=n$ in detail.
